\documentclass[12pt,a4paper]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}  % For math symbols
\usepackage{enumitem}        % For customizing lists
\usepackage{xcolor}          % For colored text
\usepackage{changepage}      % For adjusting margins
\usepackage{geometry} % Package for adjusting margins

% Command to highlight correct answer (pastel green)
\newcommand{\correctanswer}[1]{\textbf{\textcolor{green!60!black}{#1}}}

% Define a counter for questions
\newcounter{questioncounter}

% Define a new environment for MC questions
\newenvironment{mcquestion}{%
  \refstepcounter{questioncounter} % Increment question counter
  \begin{enumerate}[label=\arabic{questioncounter}., wide, labelindent=0pt]
    \item 
}{%
    \end{enumerate}
  \hrulefill  % Horizontal separation line
}

\title{BIOMEDICAL - Unit 4 and 5 MCQs}
\author{GPT-4-Turbo and Adhithya}
\date{\today}

% Adjust margins (top, bottom, left, right)
\geometry{top=1cm, bottom=2cm, left=2cm, right=2cm} 
% Start document
\begin{document}


\maketitle

% Adjust margins for wider text area
\begin{adjustwidth}{-1cm}{-1cm}

\section{UNIT-4 1 mark MCQ}


\begin{mcquestion}
    \textbf{What is the primary physical principle behind MRI?}
    \begin{enumerate}[label=\Alph*.]
      \item Piezoelectric effect
      \item X-ray attenuation
      \item \correctanswer{Nuclear magnetic resonance}
      \item Ultrasonic waves
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which component is crucial for detecting the X-rays in computed tomography (CT)?}
    \begin{enumerate}[label=\Alph*.]
      \item Cathode ray tube
      \item X-ray tube
      \item Photomultiplier tube
      \item \correctanswer{Detector array}
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In ultrasonography, what determines the resolution of the image?}
    \begin{enumerate}[label=\Alph*.]
      \item \correctanswer{Frequency of the ultrasound waves}
      \item Amplitude of the ultrasound waves
      \item Speed of sound in air
      \item Intensity of the ultrasound waves
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which imaging technique uses thermal sensors to detect radiation emitted by the body?}
    \begin{enumerate}[label=\Alph*.]
      \item MRI
      \item CT
      \item \correctanswer{Thermography}
      \item Fluoroscopy
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the main advantage of fluoroscopic imaging over traditional radiography?}
    \begin{enumerate}[label=\Alph*.]
      \item Higher resolution images
      \item Faster image processing
      \item \correctanswer{Real-time moving images}
      \item Lower radiation dose
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In endoscopy, what type of imaging technology is commonly used?}
    \begin{enumerate}[label=\Alph*.]
      \item Infrared imaging
      \item \correctanswer{Fiber optic cameras}
      \item Thermal imaging
      \item Ultrasonic imaging
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is a key feature of retinal imaging?}
    \begin{enumerate}[label=\Alph*.]
      \item It uses low-frequency sound waves.
      \item \correctanswer{It can directly visualize vascular structures.}
      \item It requires the injection of contrast media.
      \item It is primarily used for diagnosing brain disorders.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which type of biotelemetry system would be most appropriate for monitoring heart rate and rhythm?}
    \begin{enumerate}[label=\Alph*.]
      \item EEG telemetry
      \item \correctanswer{ECG telemetry}
      \item EMG telemetry
      \item Temperature telemetry
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What distinguishes a digital image analysis in medical applications?}
    \begin{enumerate}[label=\Alph*.]
      \item Use of analog signals
      \item Manual interpretation of images
      \item \correctanswer{Quantitative measurement of image features}
      \item Lower resolution images
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In biometric systems, which imaging application is commonly used for identity verification?}
    \begin{enumerate}[label=\Alph*.]
      \item Thermography
      \item \correctanswer{Retinal imaging}
      \item MRI
      \item Ultrasound
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{How does computed tomography generate images?}
    \begin{enumerate}[label=\Alph*.]
      \item By detecting reflected sound waves
      \item By recording electrical activity
      \item \correctanswer{By using rotating X-ray sources and detectors}
      \item By applying strong magnetic fields
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which imaging modality is considered safest for imaging fetuses?}
    \begin{enumerate}[label=\Alph*.]
      \item X-ray
      \item MRI
      \item \correctanswer{Ultrasound}
      \item CT
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the primary use of thermography in medical diagnostics?}
    \begin{enumerate}[label=\Alph*.]
      \item To measure bone density
      \item \correctanswer{To detect areas of heat and inflammation} 
      \item To visualize brain activity
      \item To check for arterial blockages 
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which imaging technique is most effective for visualizing soft tissues contrast?}
    \begin{enumerate}[label=\Alph*.]
      \item X-ray
      \item \correctanswer{MRI}
      \item CT 
      \item Ultrasonography 
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is a common use of fluoroscopy in clinical settings?}
    \begin{enumerate}[label=\Alph*.]
      \item Long-term imaging 
      \item Capturing static images 
      \item \correctanswer{Guiding minimally invasive surgeries} 
      \item Whole-body scans 
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What role does the piezoelectric effect play in medical imaging?}
    \begin{enumerate}[label=\Alph*.]
      \item It is used to generate X-rays in CT scanners. 
      \item It is used in MRI to align hydrogen atoms. 
      \item \correctanswer{It is used in ultrasonography to generate and detect ultrasound waves.}
      \item It is used in thermography to detect heat. 
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In digital image analysis, what is segmentation used for?}
    \begin{enumerate}[label=\Alph*.]
      \item Enhancing the contrast of the image 
      \item Reducing the noise in the image 
      \item \correctanswer{Dividing the image into meaningful regions}
      \item Compressing the image size 
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Why is MRI preferred over CT in imaging of brain tissues?}
    \begin{enumerate}[label=\Alph*.]
      \item Lower cost 
      \item Faster imaging time 
      \item \correctanswer{No ionizing radiation exposure}
      \item Higher availability 
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is an essential aspect of image processing in retinal imaging?}
    \begin{enumerate}[label=\Alph*.]
      \item \correctanswer{Motion correction}
      \item Temperature mapping 
      \item Sound wave analysis 
      \item Electrical activity recording 
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In biotelemetry, what is the primary purpose of data encryption?}
    \begin{enumerate}[label=\Alph*.]
      \item To compress the data for faster transmission 
      \item \correctanswer{To prevent unauthorized access to the data}
      \item To enhance the signal strength 
      \item To convert analog signals to digital 
    \end{enumerate}
\end{mcquestion}


\subsection{UNIT-4 2 mark MCQ}


\begin{mcquestion}
    \textbf{Which imaging modality uses a gadolinium-based contrast agent to enhance image quality?}
    \begin{enumerate}[label=\Alph*.]
      \item X-ray radiography
      \item \correctanswer{MRI}
      \item CT
      \item Ultrasound
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In computed tomography, what is the significance of the Hounsfield scale?}
    \begin{enumerate}[label=\Alph*.]
      \item It measures the intensity of the magnetic field used.
      \item \correctanswer{It is a quantitative scale for describing radiodensity.}
      \item It determines the frequency of ultrasound used.
      \item It calibrates the thermal sensors in thermography.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What differentiates a T1-weighted MRI scan from a T2-weighted scan?}
    \begin{enumerate}[label=\Alph*.]
      \item \correctanswer{T1 highlights fat better, while T2 highlights fluid.}
      \item T1 uses lower magnetic fields than T2.
      \item T1 is quicker than T2.
      \item T1 is used for bone imaging, while T2 is used for soft tissues.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In digital image processing, what is the primary purpose of using the Fourier Transform?}
    \begin{enumerate}[label=\Alph*.]
      \item To compress the images
      \item To enhance the edges of the images
      \item \correctanswer{To convert spatial domain data into frequency domain}
      \item To reduce the color depth of images
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{How does the principle of echolocation apply to ultrasonography?}
    \begin{enumerate}[label=\Alph*.]
      \item It helps in determining the position of the ultrasound device.
      \item \correctanswer{It is used to measure the depth of the body tissue based on sound reflection.}
      \item It adjusts the frequency of sound waves.
      \item It enhances the contrast of the ultrasound images.
    \end{enumerate}
\end{mcquestion}



\section{UNIT-5 1 mark MCQ}

\begin{mcquestion}
    \textbf{What is the primary function of a pacemaker?}
    \begin{enumerate}[label=\Alph*.]
      \item To monitor heart rate
      \item \correctanswer{To regulate heart rhythm}
      \item To measure blood pressure
      \item To decrease heart muscle contraction
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{How does a defibrillator work?}
    \begin{enumerate}[label=\Alph*.]
      \item By providing a steady electrical stimulus
      \item By delivering a therapeutic dose of X-ray radiation
      \item \correctanswer{By delivering a controlled electric shock to the heart}
      \item By cooling the heart muscle
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the main use of ventilators in medical settings?}
    \begin{enumerate}[label=\Alph*.]
      \item To provide pain relief
      \item To administer anesthesia
      \item \correctanswer{To assist or replace spontaneous breathing}
      \item To monitor respiratory rate
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which device is used to treat chronic pain and muscle spasms via electrical impulses?}
    \begin{enumerate}[label=\Alph*.]
      \item EEG machine
      \item Ultrasound imager
      \item \correctanswer{Nerve and muscle stimulators}
      \item Blood pressure monitor
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the purpose of diathermy in therapeutic settings?}
    \begin{enumerate}[label=\Alph*.]
      \item To measure electrical activity of the heart
      \item \correctanswer{To induce deep tissue heating}
      \item To cool body tissues
      \item To visualize internal organs
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What does a heart-lung machine do during cardiac surgeries?}
    \begin{enumerate}[label=\Alph*.]
      \item It monitors heart and lung function
      \item \correctanswer{It replaces the function of the heart and lungs}
      \item It reduces the heart rate
      \item It increases oxygen intake
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the primary function of audiometers?}
    \begin{enumerate}[label=\Alph*.]
      \item To measure visual acuity
      \item \correctanswer{To assess hearing acuity}
      \item To record electrical activity of the brain
      \item To monitor blood oxygen levels
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which device is essential for patients undergoing hemodialysis?}
    \begin{enumerate}[label=\Alph*.]
      \item Defibrillator
      \item Ventilator
      \item \correctanswer{Dialyzer}
      \item Pacemaker
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the primary benefit of using lithotripsy in treating kidney stones?}
    \begin{enumerate}[label=\Alph*.]
      \item It surgically removes stones.
      \item It dissolves stones using a chemical agent.
      \item \correctanswer{It uses shock waves to break stones into smaller pieces.}
      \item It uses high-frequency sound waves to visualize stones.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Which component is critical in an ICCU patient monitoring system?}
    \begin{enumerate}[label=\Alph*.]
      \item Anesthesia delivery unit
      \item \correctanswer{Bedside monitors displaying vital signs}
      \item Surgical instruments
      \item Medication dispensing system
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the primary application of nano robots in medicine?}
    \begin{enumerate}[label=\Alph*.]
      \item To perform large-scale surgeries
      \item To enhance imaging clarity
      \item \correctanswer{To deliver drugs specifically to diseased cells}
      \item To monitor patient dietary habits
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What distinguishes minimally invasive surgical techniques from traditional surgery?}
    \begin{enumerate}[label=\Alph*.]
      \item Longer recovery times
      \item Larger incisions
      \item \correctanswer{Less post-operative pain and scarring}
      \item Use of general anesthesia only
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{How does a defibrillator's function differ from that of a pacemaker?}
    \begin{enumerate}[label=\Alph*.]
      \item A defibrillator regulates blood pressure, while a pacemaker does not.
      \item \correctanswer{A defibrillator can restart the heart, while a pacemaker cannot.}
      \item A pacemaker uses ultrasound, while a defibrillator uses electrical shocks.
      \item A pacemaker measures oxygen saturation, while a defibrillator does not.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In what scenario might nerve and muscle stimulators be particularly useful?}
    \begin{enumerate}[label=\Alph*.]
      \item To enhance cognitive function
      \item \correctanswer{To treat muscle atrophy in immobilized patients}
      \item To improve cardiac function
      \item To stabilize body temperature
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the main therapeutic mechanism of diathermy?}
    \begin{enumerate}[label=\Alph*.]
      \item Cooling the surface tissue
      \item Inducing muscle contractions
      \item \correctanswer{Heating deeper tissues}
      \item Reducing electrical activity in the brain
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{During which procedure is a heart-lung machine most likely used?}
    \begin{enumerate}[label=\Alph*.]
      \item During kidney dialysis
      \item During non-invasive heart monitoring
      \item \correctanswer{During open-heart surgery}
      \item During a standard physical examination
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is a key feature of audiometers in audiology?}
    \begin{enumerate}[label=\Alph*.]
      \item They can directly measure ear drum resistance.
      \item \correctanswer{They provide a quantitative measure of hearing loss.}
      \item They visualize the inner ear structure.
      \item They stimulate the auditory nerve directly.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What distinguishes a dialyzer in its function from other renal support devices?}
    \begin{enumerate}[label=\Alph*.]
      \item \correctanswer{It filters blood through a semi-permeable membrane.}
      \item It injects antibiotics directly into the bloodstream.
      \item It measures the electrical activity of the kidneys.
      \item It increases blood flow to the kidneys.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is the primary function of lithotripsy in urology?}
    \begin{enumerate}[label=\Alph*.]
      \item To diagnose urinary tract infections
      \item To visualize the urinary tract
      \item \correctanswer{To break up kidney stones}
      \item To enhance bladder control
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What is an essential feature of nano robots in targeted drug delivery?}
    \begin{enumerate}[label=\Alph*.]
      \item They can store large volumes of drugs.
      \item \correctanswer{They can navigate to specific cells or tissues.}
      \item They replace traditional surgical methods.
      \item They function independently of medical supervision.
    \end{enumerate}
\end{mcquestion}


\subsection{UNIT-5 2 mark MCQ}


\begin{mcquestion}
    \textbf{What is a synchronous (demand) pacemaker and how does it differ from an asynchronous (fixed-rate) pacemaker?}
    \begin{enumerate}[label=\Alph*.]
      \item A synchronous pacemaker adjusts the pacing rate based on respiratory rate, whereas an asynchronous does not.
      \item \correctanswer{A synchronous pacemaker stimulates the heart only if it detects no natural heartbeats, whereas an asynchronous continuously stimulates the heart.}
      \item A synchronous pacemaker uses ultrasound, whereas an asynchronous uses electrical impulses.
      \item A synchronous pacemaker is implantable, whereas an asynchronous is external.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{How does an implantable cardioverter-defibrillator (ICD) differ functionally from a standard defibrillator?}
    \begin{enumerate}[label=\Alph*.]
      \item An ICD is used only during surgeries as a life-support measure.
      \item \correctanswer{An ICD can automatically detect and treat arrhythmias, while a standard defibrillator requires manual operation.}
      \item An ICD is less powerful and only used for minor corrections.
      \item An ICD monitors blood pressure, while a standard defibrillator does not.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{In what way does a modern ventilator manage different ventilation modes, such as volume-controlled and pressure-controlled?}
    \begin{enumerate}[label=\Alph*.]
      \item By changing the chemical composition of the supplied air
      \item By adjusting the speed of the fan in the ventilator
      \item \correctanswer{By altering the pressure or volume of air delivered per breath}
      \item By varying the oxygen concentration only
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{Explain the principle of operation of a dialyzer in the context of counter-current flow.}
    \begin{enumerate}[label=\Alph*.]
      \item Blood and dialysate flow in the same direction to maximize waste removal.
      \item Blood flows through a semipermeable membrane that selectively prevents large molecules from passing.
      \item \correctanswer{Blood and dialysate flow in opposite directions to increase the gradient for diffusion.}
      \item The dialyzer recycles dialysate by filtering it through a secondary system.
    \end{enumerate}
\end{mcquestion}

\begin{mcquestion}
    \textbf{What technological advancements enable nano robots to specifically target cancerous cells without affecting surrounding healthy tissue?}
    \begin{enumerate}[label=\Alph*.]
      \item Use of magnetic fields to guide them to specific locations
      \item \correctanswer{Programming with AI to identify cancer cells based on their genetic markers}
      \item Coating with materials that are only absorbed by cancerous cells
      \item Utilizing gamma radiation to selectively destroy target cells
    \end{enumerate}

\end{mcquestion}

\end{adjustwidth}
\end{document}